---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# fmdates

<!-- badges: start -->
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/fmdates)](https://cran.r-project.org/package=fmdates) [![License](https://img.shields.io/badge/license-GPL--2-blue.svg)](http://choosealicense.com/licenses/gpl-2.0/) [![GitHub Actions](https://github.com/imanuelcostigan/fmdates/workflows/R-CMD-check/badge.svg)](https://github.com/imanuelcostigan/fmdates/actions?workflow=R-CMD-check) [![Coverage Status](https://img.shields.io/codecov/c/github/imanuelcostigan/fmdates/master.svg)](https://codecov.io/github/imanuelcostigan/fmdates?branch=master)
<!-- badges: end -->

## Motivation

Over-the-counter (OTC) derivatives comprise a significant proportion of trading activity in global financial markets. Their general contractual conventions are specified in what are known as [International Swap and Derivatives Association (ISDA) definitions](https://www.isda.org/books/). For example, FX and currency option transactions are governed by the *1998 FX and Currency Options Definitions* and swap transactions are governed by the *2006 ISDA Definitions*. They describe in meticulous detail, among other things, how the dates of certain financial events should be determined. This includes how dates are defined to be good or bad and how bad dates are to be adjusted to good dates. They also define how to determine the length of time between two dates.

This package implements calendars used to define locale specific business days, date adjusters and shifters, schedule generators and year fraction calculations defined by these standards.


## Calendars

You can determine whether dates are business days in a specific locale or specific locales:

```{r is_good}
library("lubridate", warn.conflicts = FALSE)
library("fmdates")
ausy <- AUSYCalendar()
aume <- AUMECalendar()
syme <- c(ausy, aume) # handy JointCalendar construction approach
is_good(ymd(20140404), ausy)
is_good(ymd(20141104), syme) # Melbourne Cup holiday
syme$rule <- any
is_good(ymd(20141104), syme)
```

## Adjusters and shifters

You can adjust (or roll) and shift dates using predefined business day conventions:

```{r adjust_shift}
# Adjust using the modified following convention
adjust(ymd(20140404), 'mf', ausy)
# Shift dates
shift(ymd(20120229), months(1), 'u', ausy, FALSE) # one month
shift(ymd(20120229), months(1), 'mf', ausy, TRUE)  # one month with EOM rule
shift(ymd(20120229), years(1) + months(3), 'mf', ausy, TRUE)  # 1y3m
```

## Schedules

The preceding methods are used to generate schedules of dates required to define common financial contracts events such as cash flow exchange dates:

```{r schedules}
generate_schedule(effective_date = ymd(20120103), termination_date = ymd(20130103), 
  tenor = months(3), calendar = ausy, bdc = "mf", stub = "short_front", 
  eom_rule = FALSE)
```


## Year fractions

Time lengths then usually need to be computed for each interval of such a schedule according to some day basis convention:

```{r year_frac}
# 30/360us convention
year_frac(ymd("2010-03-31"), ymd("2012-03-31"), "30/360us")
# act/365 convention
year_frac(ymd("2010-02-28"), ymd("2012-03-31"), "act/365")
```

More details can be found in the associated help files and the vignette (`vignette("dates", "fmdates")`). If you would like to contribute to the package please see the [`CONTRIBUTING.md`](https://github.com/imanuelcostigan/fmdates/blob/master/CONTRIBUTING.md) file for general as well as specific suggestions.
